%! Author = t
%! Date = 2025-02-20



\newcommand{\applinks}[2][]{
    \ifthenelse{\equal{#1}{}}
        {\href{#2}{PlatMarket}}
        {\href{#2}{PlatMarket}, \href{#1}{AppStore}}% If second parameter exists, return both
}

\chapter{Mobile Apps}\label{ch:ru_apps}

It's highly recommended to install several apps on your phone in advance!
List of all recommendations with links you can find in the end of this section.


\section{VPN}\label{subsec:ru_vpn}
China is famous for its \href{https://en.wikipedia.org/wiki/Great_Firewall}{Firewall},
so you better be ready to use VPN daily!
Here's a short list of banned websites and apps:

\begin{multicols}{2}
\begin{itemize}
    \item Telegram
    \item What's App
    \item Wikipedia
    \item Google
    \item YouTube
    \item Reddit
    \item StackOverflow
    \item GitHub (not banned, but really slow)
    \item HuggingFace
\end{itemize}
\end{multicols}

\begin{enumerate}
    \item \textbf{Install N different VPNs on your phone.}
        One day your VPN might work well, but the other it won't connect at all.
        And the opposite situation might be for other VPN app.

    \item \textbf{Install Tor/TorBrowser} on your phone and laptop.
        It maybe not the fastest solution, but at least most stable option among free solutions.
        Telegram Bot for Bridges: \href{https://t.me/GetBridgesBot}{@GetBridgesBot}

    \item \textbf{Personal Server}.
        You can also configure your proxy server.
        Using the VLESS protocol is highly recommended, as other options might be banned immediately by Chinese Firewall.
\end{enumerate}




\section[Chinese Apps]{Chinese Apps}\label{sec:chinese-apps}
If you use IPhone, you can skip this Section!

Most of the following applications are Chinese,
and not all can be downloaded from the Play Market.
Therefore, you will often have to download applications from the Internet.

An alternative to this is to download the Chinese Tencent AppStore (Install it by yourown).
Of course, Chinese apps are nightmare, but it's still a solution
if you don't want to suffer with installation via the Internet.




\section[Messanger and Payments]{Messanger and Payments}\label{sec:messanger-and-payments}

\begin{itemize}
    \item \textbf{WeChat}.
    Allows message, buy tickets, order food, read news and so on.
    However, to register, a Chinese person must verify your account using a QR code.
    But still download the app in advance and try to register,
    maybe you'll be lucky enough to enter without a QR code.
    Otherwise, just postpone it until you arrive in China.

    \item \textbf{AliPay}.
    The most affordable option for paying application fee.
    In China, it mostly used among foreigners for payment, but accepted everywhere.
    First time you can use it for chatting too!
\end{itemize}




\section[Maps]{Maps}\label{sec:ru_maps}

Google is banned and Yandex Maps almost useless, but you need somehow navigate in China.


\begin{itemize}
    \item \textbf{Chinese Apps} like AMap.
        However, it supports only Chinese language.

    \item \textbf{OpenSource solutions} like \href{https://maps.me/}{Maps.Me} or \href{https://osmand.net/}{OsmAnd}.
        Help you to navigate around, but won't tell what's in mall.

    \item \textbf{MetroMan}.
        It's the most convenient app for Chinese metro.
        Again mostly developed for foreigners, but works perfect!
\end{itemize}




\section[Language]{Language}\label{sec:language}
Translator one of the most crucial tool (unless you know Chinese).
But Google is banned!
We recommend to use Yandex Translator and install Chinese Language in advance!
Also, we recommend to install vocabulary Pleco.

If you're willing to learn some basics of Chinese,
try to use HelloChinese.
It's similar to Duolingo, but helps learn Chinese better.



\section[AppList]{App List}\label{sec:ru_app_list}

\begin{itemize}
    \item WeChat -- \applinks[https://play.google.com/store/apps/details?id=com.tencent.mm]{https://apps.apple.com/us/app/wechat/id414478124}
    \item AliPay -- \applinks[https://play.google.com/store/apps/details?id=com.eg.android.AlipayGphone]{https://apps.apple.com/us/app/alipay-simplify-your-life/id333206289}
    \item Orbot -- \applinks{https://play.google.com/store/apps/details?id=org.torproject.android}
    \item AMap -- \applinks[https://play.google.com/store/apps/details?id=com.autonavi.minimap]{https://apps.apple.com/us/app/amap-global/id461703208}
    \item Maps.Me -- \applinks[https://play.google.com/store/apps/details?id=com.mapswithme.maps.pro]{https://apps.apple.com/us/app/maps-me-offline-maps-gps-nav/id510623322}
    \item MetroMan -- \applinks[https://play.google.com/store/apps/details?id=com.xinlukou.metroman]{https://apps.apple.com/us/app/metroman-china/id466351433}
    \item Pleco -- \applinks[https://play.google.com/store/apps/details?id=com.pleco.chinesesystem]{https://apps.apple.com/us/app/pleco-chinese-dictionary/id341922306}
    \item Hello Chinese -- \applinks[https://play.google.com/store/apps/details?id=com.hellochinese]{https://apps.apple.com/us/app/hellochinese-learn-chinese/id1001507516}
    \item TaoBao -- \applinks[https://play.google.com/store/apps/details?id=com.taobao.taobao]{https://apps.apple.com/us/app/taobao/id387682726}
\end{itemize}
